Existence of Forward Simulations for Particular Data Structures

### 2017-01-14 draftCAV/app-absImplStack.tex
a pass on abstract implementations

%!TEX root = draft.tex
\section{Stacks With Fixed Pop Commit Points}

Differently from queues, there are implementations of concurrent stacks where the linearization points of the remove/pop operations are not fixed, e.g., Time-stamped Stack~\cite{}. We define an abstract implementation which simulates such implementations provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action. To demonstrate the use of this abstract implementation we provide a correctness proof for a simplified version of Time-Stamped Stack that preserves its most complex behaviors.

\subsection{Abstract stack implementation}

Let $Com(pop)$ be the set of actions $com(pop,d,k)$ with $d\in\<Vals>$ and $k\in\<Ops>$, called \emph{commit points}, representing actions of a pop operation where its return value is determined. We define an abstract implementation $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed. Since the commit points are not necessarily linearization points, the pop operations are treated differently. Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order and continuously tracks the push operations which are overlapping with it. The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.

The states of $AbsS$ are tuples $\tup{O,<,\ell,rval,cp,comp,pend}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked push operations), $<\subseteq O\times O$ is a strict partial order over $O$, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a flag that records whether the operation is pending or completed, $coPush:\<Ops> ~> 2^O$ records push operations that were maximal when a pop started or happening earlier provided that the values of all the push happening later have been removed, and $pePush: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.

\begin{figure} [t]
{\footnotesize
  \centering
  \begin{mathpar}
    \inferrule[call-push]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rval,cp,comp,pend 
      \xrightarrow{inv(push,d,k)} \\
      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND},\emptyset)],rval,cp[k\mapsto 1],comp,pend[\forall k'.\ k'\mapsto pend(k')\cup \{k\}]
    }\hspace{5mm}

    \inferrule[call-pop]{
      k\not\in dom(cp) \\
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{inv(pop,k)} \\
      O,<,\ell,rval,cp[k\mapsto 1],comp[k\mapsto maxCompl(<)],pend[k\mapsto \{k':\ell_2(k')={\tt PEND}\}]
    }\hspace{5mm}

    \inferrule[ret-pop]{
       cp(k) = 2 \\
       rval(k)=d  
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{ret(pop,d,k)}
      O,<,\ell,rval,cp[k\mapsto 3],comp,pend
    }\hspace{5mm}

    \inferrule[ret-push1]{
      cp(k) = 1 \\
      k \in O \\
      \ell(k) = (d,{\tt PEND},X) 
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{ret(push,k)}
      O,<,\ell[k\mapsto (d,{\tt COMP},X)],rval,cp[k\mapsto 2],comp,pend
    }\hspace{5mm}

    \inferrule[ret-push2]{
      cp(k) = 1 \\
      k \not\in O 
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{ret(push,k)}
      O,<,\ell,rval,cp[k\mapsto 2],comp,pend
    }\hspace{5mm}

    \inferrule[lin-pop1]{
       cp(k) = 1 \\
       d\neq{\tt EMPTY} \\
       \exists k'\in O.\ k'\in comp(k)\cup pend(k)\land \ell_1(k')=d
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{lin(pop,d,k)} \\
      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2],comp[\forall k''.\ k'\in comp(k'') \Rightarrow k'' \mapsto (comp(k'')\cup pred_{<}(k'))\setminus\{k'\}],pend
    }\hspace{5mm}

    \inferrule[lin-deq2]{
       cp(k) = 1 \\
       comp(k)=\emptyset
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{lin(deq,{\tt EMPTY},k)}
      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2],comp,pend
    }\hspace{5mm}

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$.}
  \label{fig:transitions:AbsS}
\vspace{-6mm}
\end{figure}

\subsection{Other}

>>>>

%!TEX root = draft.tex
\section{Proof of Theorem~\ref{th:absImplStack}}\label{app:absImplStack}

### 2017-01-14 draftCAV/prelim.tex
a pass on abstract implementations

>>>>

A library $L$ is called a \emph{queue implementation} when its alphabet contains call and return actions defined over $\<Methods>=\{enq,deq\}$ ($enq$ is the name of the methods that enqueue a value, and $deq$ is the name of the methods that dequeue a value) and $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$ where {\tt EMPTY} is the value returned by $deq$ when the queue is empty. Similarly, a library $L$ is called a \emph{stack implementation} when $\<Methods>=\{push,pop\}$ and  $\<Vals>=\<Nats>\cup\{{\tt EMPTY}\}$.

### 2017-01-14 draftCAV/queue.tex
a pass on abstract implementations

%!TEX root = draft.tex
\section{Queues With Fixed Dequeue Linearization Points}

The typical abstract implementation of a concurrent queue maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns ``empty''). Both methods have a fixed linearization point when the update of the shared sequence happens. Let $AbsQ_0$ denote this implementation (formally defined in Appendix~\ref{app:absImplQueue}). In this section, we describe another abstract implementation, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence. We show that there exists a forward simulation from any correct queue implementation where the dequeue methods have fixed linearization points to $AbsQ$. We are not aware of any implementation that doesn't satisfy this property. As a proof of concept, we describe a forward simulation from the Herlihy\&Wing queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.

@@

The states of $AbsQ$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed, $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.

\begin{figure} [t]
{\footnotesize
  \centering
  \begin{mathpar}
    \inferrule[call-enq]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rval,cp
      \xrightarrow{inv(enq,d,k)}
      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rval,cp[k\mapsto 1]
    }\hspace{5mm}

    \inferrule[call-deq]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rval,cp
      \xrightarrow{inv(deq,k)}
      O,<,\ell,rval,cp[k\mapsto 1]
    }\hspace{5mm}
    \inferrule[ret-deq]{
       cp(k) = 2 \\
       rval(k)=d  
    }{
      O,<,\ell,rval,cp
      \xrightarrow{ret(deq,d,k)}
      O,<,\ell,rval,cp[k\mapsto 3]
    }\hspace{5mm}

    \inferrule[ret-enq1]{
      cp(k) = 1 \\
      k \in O \\
      \ell(k) = (d,{\tt PEND}) 
    }{
      O,<,\ell,rval,cp
      \xrightarrow{ret(enq,k)}
      O,<,\ell[k\mapsto (d,{\tt COMP})],rval,cp[k\mapsto 2]
    }\hspace{5mm}
    \inferrule[ret-enq2]{
      cp(k) = 1 \\
      k \not\in O 
    }{
      O,<,\ell,rval,cp
      \xrightarrow{ret(enq,k)}
      O,<,\ell,rval,cp[k\mapsto 2]
    }\hspace{5mm}

    \inferrule[lin-deq1]{
       cp(k) = 1 \\
       d\neq{\tt EMPTY} \\
       \exists k'\in O.\ k'\in min(<) \land \ell_1(k')=d
    }{
      O,<,\ell,rval,cp
      \xrightarrow{lin(deq,d,k)}
      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2]
    }\hspace{5mm}

    \inferrule[lin-deq2]{
       cp(k) = 1 \\
       O=\emptyset
    }{
      O,<,\ell,rval,cp
      \xrightarrow{lin(deq,{\tt EMPTY},k)}
      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
    }\hspace{5mm}

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$.}
  \label{fig:transitions:AbsQ}
\vspace{-6mm}
\end{figure}

@@

A library $L$ is called a \emph{queue implementation} when its alphabet consists of call and return actions defined over $\<Methods>=\{enq,deq\}$. Let $LP(deq)$ be the set of all actions $lin(deq,d,k)$ for some $d\in\<Vals>$ and $k\in\<Ops>$. A trace of a queue implementation is called $LP(deq)$-complete when every completed dequeue has a linearization point, i.e., TODO. A queue implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed dequeue linearization points} if{f} $C\cup R\cup LP(deq)\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $LP(deq)$-complete.

@@

The following result shows that $C\cup R\cup LP(deq)$-forward simulations are a sound and complete proof method for showing the correctness of a queue implementation with fixed dequeue linearization points (up to the correctness of the linearization points). It is obtained from Theorem~\ref{th:absImplQueue} and Theorem~\ref{th:forSim} using the fact that the alphabet of $AbsQ$ is exactly $C\cup R\cup LP(deq)$ and $AbsQ$ is deterministic.

\begin{corollary}
A queue implementation $L$ with fixed dequeue linearization points is a $C\cup R\cup LP(deq)$-refinement of $AbsQ_0$ if{f} there exists a $C\cup R\cup LP(deq)$-forward simulation from $L$ to $AbsQ$.
\end{corollary}

>>>>

%!TEX root = draft.tex
\section{Queues With Fixed Dequeue Linearization Points}\label{sec:queues}

The typical abstract implementation of a concurrent queue maintains a sequence of values, the enqueue method adds a value atomically to the beginning of the sequence, and the dequeue method removes a value from the end of the sequence (if any, otherwise it returns ``empty''). Both methods have a fixed linearization point when the update of the shared sequence happens. Let $AbsQ_0$ denote this implementation~\footnote{For $\<Methods>=\{enq,deq\}$, the alphabet of $AbsQ_0$ is $C\cup R\cup Lin$.} (formally defined in Appendix~\ref{app:absImplQueue}). In this section, we describe another abstract implementation, denoted as $AbsQ$, which roughly maintains a \emph{partially-ordered set} of values instead of a sequence. We show that there exists a forward simulation from any correct queue implementation where the \emph{dequeue} methods have fixed linearization points to $AbsQ$. We are not aware of any implementation that doesn't satisfy this property. As a proof of concept, we describe a forward simulation from the Herlihy\&Wing queue~\cite{journals/toplas/HerlihyW90} to $AbsQ$.

@@

TODO DESCRIBE THE HERLIHY \& WING, EXPLAIN THAT ENQUEUES HAVE NO FIXED LINEARIZATION POINTS, GIVE THE LINEARIZATION POINTS OF DEQUEUES, SHOW HOW $AbsQ$ CAN SIMULATE SOME PROBLEMATIC HISTORIES

The states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed (this flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}. The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points. The latter are denoted by $lin(deq,d,k)$ where $d\in \<Vals>$ and $k\in\<Ops>$. $Lin(deq)$ is the set of all such actions.

Concerning enqueue operations, the rule {\sc call-enq} orders the invoked operation after all the completed enqueue operations stored in the current state, and the rules {\sc ret-enq1}/{\sc ret-enq2} flip the corresponding flag from {\tt PEND} to {\tt COMP} provided that the operation is still present in the current state. For dequeue operations, {\sc call-deq} has no effect other than incrementing the control point and {\sc ret-deq} checks whether the return value is the same as the one fixed at the linearization point. The linearization point rule {\sc lin-deq1} corresponds to the case of a non-empty queue, showing that $lin(deq,d,k)$ is enabled only if the value $d$ has been added by an enqueue which is minimal in the current happens-before order. When it is enabled, it removes the enqueue adding $d$ from the state. The linearization point rule {\sc lin-deq2} corresponds to the case of dequeue operations linearized when the queue is empty.

\begin{figure} [t]
{\scriptsize
  \centering
  \begin{mathpar}
    \inferrule[call-enq]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rv,cp
      \xrightarrow{inv(enq,d,k)}
      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
    }\hspace{5mm}

    \inferrule[call-deq]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rv,cp
      \xrightarrow{inv(deq,k)}
      O,<,\ell,rv,cp[k\mapsto 1]
    }\hspace{5mm}
    \inferrule[ret-deq]{
       cp(k) = 2 \\
       rv(k)=d  
    }{
      O,<,\ell,rv,cp
      \xrightarrow{ret(deq,d,k)}
      O,<,\ell,rv,cp[k\mapsto 3]
    }\hspace{5mm}

    \inferrule[ret-enq1]{
      cp(k) = 1 \\
      k \in O \\
      \ell(k) = (d,{\tt PEND}) 
    }{
      O,<,\ell,rv,cp
      \xrightarrow{ret(enq,k)}
      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto 2]
    }\hspace{5mm}
    \inferrule[ret-enq2]{
      cp(k) = 1 \\
      k \not\in O 
    }{
      O,<,\ell,rv,cp
      \xrightarrow{ret(enq,k)}
      O,<,\ell,rv,cp[k\mapsto 2]
    }\hspace{5mm}

    \inferrule[lin-deq1]{
       cp(k) = 1 \\
       d\neq{\tt EMPTY} \\
       k'\in min(O) \land \ell_1(k')=d
    }{
      O,<,\ell,rv,cp
      \xrightarrow{lin(deq,d,k)}
      O\setminus \{k\},<\uparrow k,\ell,rv[k\mapsto d],cp[k\mapsto 2]
    }\hspace{5mm}

    \inferrule[lin-deq2]{
       cp(k) = 1 \\
       O=\emptyset
    }{
      O,<,\ell,rv,cp
      \xrightarrow{lin(deq,{\tt EMPTY},k)}
      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto 2]
    }\hspace{5mm}

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$. We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.}
  \label{fig:transitions:AbsQ}
\vspace{-6mm}
\end{figure}

@@

A trace of a queue implementation is called \emph{$Lin(deq)$-complete} when every completed dequeue has a linearization point, i.e., TODO. A queue implementation $L$ over alphabet $\Sigma$ is called \emph{with fixed dequeue linearization points} if{f} $C\cup R\cup Lin(deq)\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $Lin(deq)$-complete.

@@

The following result shows that $C\cup R\cup Lin(deq)$-forward simulations are a sound and complete proof method for showing the correctness of a queue implementation with fixed dequeue linearization points (up to the correctness of the linearization points). It is obtained from Theorem~\ref{th:absImplQueue} and Theorem~\ref{th:forSim} using the fact that the alphabet of $AbsQ$ is exactly $C\cup R\cup Lin(deq)$ and $AbsQ$ is deterministic.

\begin{corollary}
A queue implementation $L$ with fixed dequeue linearization points is a $C\cup R\cup Lin(deq)$-refinement of $AbsQ_0$ if{f} there exists a $C\cup R\cup Lin(deq)$-forward simulation from $L$ to $AbsQ$.
\end{corollary}

### 2017-01-14 draftCAV/refProofs.tex
a pass on abstract implementations

The linearization of a concurrent history can be also defined in terms of \emph{linearization points}. Informally, a linearization point of 
an operation in an execution is a point in time where the operation is conceptually effectuated; given the linearization points of 
each operation, the linearization of a concurrent history is the sequential history which takes operations in order of their linearization points.
For some libraries, the linearization points correspond to a fixed set of actions. For instance, in the case of atomic libraries  
where method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution 
of the body. When the linearization points are fixed, we assume that the library is an LTS over an alphabet that includes actions 
$lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$. The action $lin(m,d,k)$ represents the linearization point of the operation $k$ 
which is an invocation of method $m$ returning value $d$.
Let $LP$ denote the set of such actions. We extend the notion of
well-formedness to sequences $@t$ over alphabet $\Sigma$,
such that $C\cup R\cup LP\subseteq \Sigma$, by adding the following constraints:
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
%The fact that a history $h_1$ is linearizable w.r.t. another sequential history $h_2$ can be stated as: there exists a way of choosing a linearization
%point 
%Formally, a \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by 
%inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
%\begin{itemize}
%	\item every linearization point corresponds to an operation that has been called, i.e., TODO
%	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
%	\item the return value of an operation is fixed by the linearization point, i.e., TODO
%\end{itemize}
We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup LP$ is called an 
\emph{extended history}. A trace or extended history is called \emph{LP-complete} when every completed operation has a linearization 
point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup LP\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $LP$-complete.

Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
reduces to proving that $L_1$ is a $(C\cup R\cup LP)$-refinement of an abstract implementation $L_2$ of the same object where method
bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is deterministic, which is typically the case,
by Theorem~\ref{th:forSim}, proving $(C\cup R\cup LP)$-refinement is equivalent to finding a $(C\cup R\cup LP)$-forward simulation 
from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup LP$.}.

>>>>

The linearization of a concurrent history can be also defined in terms of \emph{linearization points}. Informally, a linearization point of 
an operation in an execution is a point in time where the operation is conceptually effectuated; given the linearization points of 
each operation, the linearization of a concurrent history is the sequential history which takes operations in order of their linearization points.
For some libraries, the linearization points correspond to a fixed set of actions. For instance, in the case of atomic libraries  
where method bodies are guarded with a global-lock acquisition, the linearization point of every method invocation corresponds to the execution 
of the body. When the linearization points are fixed, we assume that the library is an LTS over an alphabet that includes actions 
$lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$. The action $lin(m,d,k)$ represents the linearization point of the operation $k$ 
which is an invocation of method $m$ returning value $d$.
Let $Lin$ denote the set of such actions. We extend the notion of
well-formedness to sequences $@t$ over alphabet $\Sigma$,
such that $C\cup R\cup Lin\subseteq \Sigma$, by adding the following constraints:
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
%The fact that a history $h_1$ is linearizable w.r.t. another sequential history $h_2$ can be stated as: there exists a way of choosing a linearization
%point 
%Formally, a \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by 
%inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
%\begin{itemize}
%	\item every linearization point corresponds to an operation that has been called, i.e., TODO
%	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
%	\item the return value of an operation is fixed by the linearization point, i.e., TODO
%\end{itemize}
We assume that libraries produce only well-formed traces. The projection of a library trace over $C\cup R\cup Lin$ is called an 
\emph{extended history}. A trace or extended history is called \emph{$Lin$-complete} when every completed operation has a linearization 
point, i.e., TODO. A library $L$ over alphabet $\Sigma$ is called \emph{with fixed linearization points} if{f} $C\cup R\cup Lin\subseteq \Sigma$ 
and every trace $@t\in Tr(L)$ is $Lin$-complete.

Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
bodies are guarded with a global-lock acquisition. Assuming that the abstract implementation is deterministic, which is typically the case,
by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
from $L_1$ to $L_2$~\footnote{Note that the alphabet of $L_2$ is exactly $C\cup R\cup Lin$.}.
